---
title: "temperature_prediction"
author: "Daniel Pienica"
date: "4/2/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("automl")
#install.packages("summarytools")
#install.packages("dlookr")
#install.packages("datamaid")
#install.packages("visdat")
#install.packages("DataExplorer")
library(data.table)
library(DataExplorer)
library(magrittr)


```
```{r load_data}
train_data<-fread('project_data/train_data.csv')

```
## EDA

* Turn relevant features to factorial 
* prcp looks has a lot of Na's 
```{r}
head(train_data[is.na(prcp),,])
dim(train_data[prcp ==0])
dim(train_data[is.na(prcp),,])
```


```{r}
plot_intro(train_data)

```
Only 7.4% of all the rows are complete!
Only 4.8% of total missing observations 

```{r}
plot_missing(train_data)
```
We have two columns with a huge lot of missing values
prcp - 85.9% Amount of precipitation in millimetres (last hour)
gbrd - 42.05% Solar radiation KJ/m2

There could be a big case to drop prcp. We will have to try to investigate the source of the missing values

```{r}
options(scipen = 999)
plot_bar(train_data)
```

Some categorical values were not explored because of high value of categories, we will have to see if they are really categorical etc.
```{r}
library(magrittr)
train_data[,1:10,] %>% head()
hist(train_data[,prcp,])
hist(train_data[,gbrd,])
```
*prcp needs to go , most of the values are 0 and NA. drop it 
*gbrd also has a very high percentage of zeroes. But we will have to be more careful with it.
```{r}
train_data <- train_data[,!"prcp"]

```
```{r}
par(mfrow = c(4,1))
plot_histogram(train_data[sample(.N,100000)])
```
We can see a couple of interesting things:
- da has a lot of values around 12
- dewp is kind of skewed. a lot of values have 0
- dmax the same
- gbrd has a lot of 0 values, let's remember that it also had 42% of missing values 
- temp looks very gaussian except for the 0 values
- concerning the years most of the vakues are frin 2008 and onward. we migh try to see if there is a link between the 0 and the years
- hours seem evenly distributed
- wsid clearly has some outliers
- stp, smax smin a lot of 0 values
- There seems to be a lot of zeroes. I need to see if the rows that have zeros in one of the values also have in the others.
Let's check this fast 

```{r}
plot_histogram(train_data[hmdy == 0][sample(.N,100000)])
```
Ok wow, great news. We can see that all the zero values seem to be correlated
it shows that from the 100000 rows we selected with hmin is 0 , we have a high number of zeros for the following columns
dewp,gust,wdsp, hmin ., dewp, stp

I will plot correlations once with the zeroes and once without
```{r}
plot_correlation(na.omit(train_data[sample(.N,100000)]))
```
temp is our y- value
* it has some pretty negative correlation with elvt (which makes sense)
* pretty strong cortrelation with hmdy,hmin,hmax, dewp,dmax,dmin - this is explained because of the zeroes ( see below)

```{r}
plot_correlation(na.omit(train_data[hmdy!=0 | gust!=0  | wdsp !=0][sample(.N,100000)]))
```

Without the zeros it all changes...

negative correlation with hmdy, hmin and hmax (makes sense)

```{r}
plot_correlation(na.omit(train_data[hmdy==0 ][sample(.N,100000)]))
```

```{r}
plot_prcomp(na.omit(train_data[,!"gbrd"][sample(.N,100000)]),ncol = 2,nrow =1)

```

* temp has some missing values - check and remove them
* check if dupicate values - no duplicates
```{r}
plot_boxplot(train_data[hmdy!=0 | gust !=0][sample(.N,100000)],by= "temp")

```
wsid    wsnm    elvt     lat     lon    inme    city    prov    mdct    date      yr      mo      da      hr    prcp     stp 
    105     105      99     104     104     104     102       4  140916    5974      17      12      31      24     347    2324 
   smax    smin    gbrd    temp    dewp    dmax    dmin    hmdy    hmax    hmin    wdsp    wdct    gust 
   2324    2323 1912249     460     463     491     436      92     102     102     182     361     335 

```{r}
plot_scatterplot(train_data[hmdy!=0 | gust !=0][sample(.N,10000)],by= "temp",nrow=2, ncol=1)

```






```{r }
class(train_data)
head(train_data)
train_data[,city,] %>% table
```


```{r eda}
summary(train_data)
str(train_data)

```
```{r}
train_data[,temp,]%>% hist()
train_data[,dewp,]%>% hist()
head(train_data[dewp<0,,])
```
The distribution lookes gaussian
Interesting.
We can see some negative values though
### remove columns with too much NA's



```{r }

#options(scipen=999)
#dim(train_data)
#calc_na_percentage <- function(col) sum(is.na(col))/length(col)
#NaPercentageByColumn<-t(train_data[,lapply(.SD,calc_na_percentage),.SD = names(train_data)])
#NaPercentageByColumn

#library(ggplot2)
#install.packages("GGally")
#library(GGally)
#ggcorr(train_data, label = TRUE, label_alpha = TRUE)

```

```{r}
#par(mar = rep(2,4))
#ggpairs(train_data)
```
